Performing Cluster Analysis on TAD boundaries & Annotations

### 2019-03-22 Scripts/evaluating_pca.Rmd
editted script for pca

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(factoextra)
library(FactoMineR)

@@

```{r}
pcadat_gm12878_10kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_count.rds")
dim(pcadat_gm12878_10kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_10kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_10kb_count_f <- pcadat_gm12878_10kb_count[, -which(colnames(pcadat_gm12878_10kb_count) %in% nzvar)]

dim(pcadat_gm12878_10kb_count) 
dim(pcadat_gm12878_10kb_count_f)

colnames(pcadat_gm12878_10kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_10kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_10kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_10kb_count_f),
          #rowsep=1:nrow(pcadat_gm12878_10kb_count_f),
          keysize = 1)

#PCA
pcadat_gm12878_10kb_count_f <- t(pcadat_gm12878_10kb_count_f)

Group <- ifelse(rownames(pcadat_gm12878_10kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_10kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_10kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_10kb_count_f <- cbind.data.frame(pcadat_gm12878_10kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_f[,-(dim(pcadat_gm12878_10kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_f[,-(dim(pcadat_gm12878_10kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_count_f[,-(dim(pcadat_gm12878_10kb_count_f2)[2])], 4),
         data=pcadat_gm12878_10kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

>>>>

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(factoextra)
library(FactoMineR)
library(ggfortify)
```

@@

```{r}
pcadat_gm12878_10kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_count.rds")
pcadat_gm12878_10kb_perc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_perc.rds")
pcadat_gm12878_10kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_dist.rds")

#performing log2 transformation on distances
pcadat_gm12878_10kb_dist <- apply(pcadat_gm12878_10kb_dist,2,function(x){log(x + 1, base=2)})

cleancolnames <- function(x){gsub(paste(c("`", 
                                          "_",
                                          #"-",
                                          #"Gm12878", 
                                          #"Haib", 
                                          #"Sydh",
                                          #"Uta",
                                          #"Uw",
                                          #"Broad",
                                          #"Uchicago",
                                          "dist",
                                          "count",
                                          "perc"), collapse = "|"),"", 
                                  colnames(x))}

colnames(pcadat_gm12878_10kb_count) <- cleancolnames(pcadat_gm12878_10kb_count)
colnames(pcadat_gm12878_10kb_perc) <- cleancolnames(pcadat_gm12878_10kb_perc)
colnames(pcadat_gm12878_10kb_dist) <- cleancolnames(pcadat_gm12878_10kb_dist)

pcadat_gm12878_10kb_count <- t(pcadat_gm12878_10kb_count)
pcadat_gm12878_10kb_perc <- t(pcadat_gm12878_10kb_perc)
pcadat_gm12878_10kb_dist <- t(pcadat_gm12878_10kb_dist)

#pcadat_gm12878_10kb_count <- data.frame(pcadat_gm12878_10kb_count)
#pcadat_gm12878_10kb_perc <- data.frame(pcadat_gm12878_10kb_perc)
#pcadat_gm12878_10kb_dist <- data.frame(pcadat_gm12878_10kb_dist)

#pcadat_gm12878_10kb_count$Type <- "Counts"
#pcadat_gm12878_10kb_perc$Type <- "Percent"
#pcadat_gm12878_10kb_dist$Type <- "Distance"

histonevars <- gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/HistoneModifications"))
tfbsvars <-  gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/gm12878"))

#### Chromatin States

```{r}
chromstatevars <- gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/BroadHMM"))
```

##### Overlap Counts

```{r}
pcadat_gm12878_10kb_count <- pcadat_gm12878_10kb_count[which(rownames(pcadat_gm12878_10kb_count) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_count),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_count),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_count)-1)*sum(apply(pcadat_gm12878_10kb_count,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_count,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_count, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_count,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()
